Probing the inner kpc of massive galaxies with strong gravitational lensing

### 2015-01-07 ms.tex
final revisions for resubmission

\noindent We examine the prospects of detecting demagnified images 
of gravitational lenses in observations of strongly lensed mm-wave 
molecular emission lines with ALMA. We model the lensing galaxies 
as a superposition of a dark matter component, a stellar component, 
and a central supermassive black hole and assess the detectability 
of the central images for a range of relevant parameters (e.g., 
stellar core and black hole mass). We find that over a large range 
of plausible parameters, future deep observations of lensed 
molecular lines with ALMA should enable detection of the central 
images at $\gtrsim 3\sigma$ significance. We use a Fisher analysis 
to examine the constraints that could be placed on these parameters 
in various scenarios and find that for large stellar cores, both 
the core size and the mass of the central SMBHs can be accurately 
measured. We also study the prospects for detecting binary SMBHs 
with such observations and find that only under rare conditions and 
with very long integrations ($\sim$40-hr) the masses of both SMBHs 
may be measured using the distortions of central images.

@@

We modeled the lens potential as a sum of three components: dark matter, stellar population and a single central SMBH. The dark matter component was modeled as \blue{an elliptical} singular power-law with a slope of 0.1, in agreement with the projected surface mass density of the NFW profile in the innermost regions \citep{Golse:02}. As  pointed out by \citet{Keeton:03}, due to its expected flatness in these regimes, the dark matter component has negligible influence on the central images (although models with dark matter cusps have been entertained). 
The stellar population is modeled as \blue{an elliptical} cored power-law, $\Sigma(r)\propto (r^2+R_{\mathrm{core}}^2)^{\gamma_s}$. \blue{The deflection angles for elliptical cored power-law models were calculated using FastEll \citep{Barkana:98}.}  Figure \ref{fig2} shows the stellar component (red curve), dark matter component (black dashed curve) and the sum of the two (black solid curve). The grey dashed curves show a few examples of core-sersic models with parameters taken from \citet{Ferrarese:06}. Although the cored power-law model used in this work does not account for the slope of the stellar distribution below the core break, it is a close fit and a reasonable model to approximate the observed stellar light profiles in the inner few kpc. 
The black hole was modeled as a simple point mass at the center of the potential. 
\blue{Lensing galaxies are typically massive ellipticals with velocity dispersions of order 200 km/s \citep[e.g.,][]{Sonnenfeld:13}. From the M-$\sigma$ relation, we predict that these galaxies should contain black holes with masses in excess of $2\times10^8 M_{\odot}$ \citep[e.g.,][]{Kormendy:13}. In our simulations we have chosen black holes with masses of 1, 2, and $4 \times 10^8 M_{\odot}$.}

>>>>

\noindent We examine the prospects of detecting demagnified images 
of gravitational lenses in observations of strongly lensed mm-wave 
molecular emission lines with ALMA. We model the lensing galaxies 
as a superposition of a dark matter component, a stellar component, 
and a central supermassive black hole and assess the detectability 
of the central images for a range of relevant parameters (e.g., 
stellar core, black hole mass, and source size). 
We find that over a large range 
of plausible parameters, future deep observations of lensed 
molecular lines with ALMA should enable detection of the central 
images at $\gtrsim 3\sigma$ significance. 
We use a Fisher analysis 
to examine the constraints that could be placed on these parameters 
in various scenarios and find that for large stellar cores, both 
the core size and the mass of the central SMBHs can be accurately 
measured. 
We also study the prospects for detecting binary SMBHs 
with such observations and find that only under rare conditions and 
with very long integrations ($\sim$40-hr) the masses of both SMBHs 
may be measured using the distortions of central images.

@@

We modeled the lens potential as a sum of three components: dark matter, stellar population and a single central SMBH. The dark matter component was modeled as \blue{an elliptical} singular power-law with a slope of 0.1, in agreement with the projected surface mass density of the NFW profile in the innermost regions \citep{Golse:02}. As  pointed out by \citet{Keeton:03}, due to its expected flatness in these regimes, the dark matter component has negligible influence on the central images (although models with dark matter cusps have been entertained). 
The stellar population is modeled as \blue{an elliptical} cored power-law, $\Sigma(r)\propto (r^2+R_{\mathrm{core}}^2)^{\gamma_s}$. \blue{The deflection angles for elliptical cored power-law models were calculated using FASTELL \citep{Barkana:98}.}  Figure \ref{fig2} shows the stellar component (red curve), dark matter component (black dashed curve) and the sum of the two (black solid curve). The grey dashed curves show a few examples of core-sersic models with parameters taken from \citet{Ferrarese:06}. Although the cored power-law model used in this work does not account for the slope of the stellar distribution below the core break, it is a close fit and a reasonable model to approximate the observed stellar light profiles in the inner few kpc. 
The black hole was modeled as a simple point mass at the center of the potential. 
\blue{Lensing galaxies are typically massive ellipticals with velocity dispersions of order 200 km/s \citep[e.g.,][]{Sonnenfeld:13}. From the M-$\sigma$ relation, we predict that these galaxies should contain black holes with masses in excess of $2\times10^8 M_{\odot}$ \citep[e.g.,][]{Kormendy:13}. In our simulations we have chosen black holes with masses of 1, 2, and $4 \times 10^8 M_{\odot}$.}
